Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2025-07-09 analysis/paper/article.qmd
figures at a good size in the docx now

---
# title: "Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology"
# author:
#   - Ben Marwick:
#       correspondence: "yes"
#       email: bmarwick@uw.edu
#       orcid: 0000-0001-7879-4531
#       institute:
#         - uw
#   - Anne Marie Poole:
#       institute: uw
#       orcid: xxx
#   - Ailin Zhang:
#       institute: uw
#       orcid: xxx
#   - Setareh Shafizadeh:
#       orcid: 0000-0003-2231-4310
#       institute: uw
#   - Jess Beck:
#       orcid: 0000-0002-7387-2307
#       institute:
#         - ucd
# institute:
#   - uw: Department of Anthropology, University of Washington, Seattle, USA
#   - ucd: School of Archaeology, University College Dublin, Ireland
#   
# title-block-published: "Last updated"  
# date: now
# date-format: long
# ---  Output Options ---
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 1000
filters:
 # - ../templates/scholarly-metadata.lua
 # - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/society-for-american-archaeology.csl" # Insert path for the bib-style
reference-section-title: "References Cited" # Required title for the reference list
---

@@

```{r}
#| label: fig-panel-wiki-self-report-numbers
#| fig-cap: "Number of users self-reporting in each category of 'current users' per year"
#| fig-width: 3

@@

how_many_people_use_the_wiki <- 
tidy_table_to_plot %>% 
  filter(status != "How many people use the wiki?") %>% 
  mutate(status = str_wrap(status, 26)) %>% 
ggplot() +
  aes(year,
      count,
      fill = status) +
  geom_col() +
  theme_minimal(base_size = small_base_size) +
  xlab("") +
  scale_fill_manual(values = the_thing_palette) +
  theme(legend.position = c(.7,.8),
        legend.key.size = unit(0.25,"line"),
        legend.text = element_text(margin = margin(r = 10, 
                                                   l = 2,
                                                   t = 2,
                                                   b = 2,
                                                   unit = "pt"))) +
  guides(fill=guide_legend(nrow=3,
                           byrow=TRUE),
         )
```

```{r}
#| label: fig-panel-wiki-characteristics
#| fig-cap: "A: Total number of edits to each Wiki page (additions and deletions) per year. B: Number of unique editors, as represented by distinct usernames or IP addresses (recorded for editors without usernames), active per year. C: Active periods of individual editors, the black lines connect the date of the first and last edit of all editors that were active for three months or more. D: Typical size of each edit per year, either addition or removal of text, one word is roughly 4-6 bytes. E: Distribution of the numbers of edits per job ad for each year. F: Histogram showing that the majority of editors are only active on the site for less than one year. G: Breakdown of self-reported Wiki user categories for each year."
#| out-width: "3in"
#| out-height: "2in"

@@

```{r}
#| label: fig-show-basic-plots
#| fig-cap: "A: total number of job ads from US institutions posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank. B: Ratio of tenure-track to non-tenure-track positions over time. The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
#| out-width: "2in"
#| out-height: "2in"

@@

```{r}
#| label: fig-show-map-of-hiring-institution-and-cc
#| fig-cap: "A: Frequency of hiring institution by Carnegie classification. B: Inset shows map of the United States showing the count of tenure-track job ads posted by all institutions in each state during 2013 -- 2023"
#| fig-width: 3

@@

```{r}
#| label: fig-show-geo-trends
#| fig-cap: "A: Frequency of locations mentioned in the text of the job ads. B: Popularity of locations in job ads over time for locations that appear in 20 or more ads. Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
#| fig-width: 3

@@

```{r}
#| label: fig-show-metho-trends
#| fig-cap: "A: Frequency of methods mentioned in the text of the job ads. B: Popularity of methods in job ads over time for methods that appear in 10 or more ads. Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
#| fig-width: 3

@@

```{r}
#| label: fig-show-topi-trends
#| fig-cap: "A: Frequency of topics mentioned in the text of the job ads. B: Popularity of topics in job ads over time for topics that appear in 20 or more ads. Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
#| fig-width: 3

@@

```{r}
#| label: fig-show-cooc
#| fig-cap: "Heatmap of topic co-occurrence in job ads."
#| fig-width: 3

@@

```{r}
#| label: fig-requirements-over-time
#| fig-cap: "Changing requirements in job ads over time. 'Count' refers to the number of each item required."
#| fig-width: 3

>>>>

---
# title: "Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology"
# author:
#   - Ben Marwick:
#       correspondence: "yes"
#       email: bmarwick@uw.edu
#       orcid: 0000-0001-7879-4531
#       institute:
#         - uw
#   - Anne Marie Poole:
#       institute: uw
#       orcid: xxx
#   - Ailin Zhang:
#       institute: uw
#       orcid: xxx
#   - Setareh Shafizadeh:
#       orcid: 0000-0003-2231-4310
#       institute: uw
#   - Jess Beck:
#       orcid: 0000-0002-7387-2307
#       institute:
#         - ucd
# institute:
#   - uw: Department of Anthropology, University of Washington, Seattle, USA
#   - ucd: School of Archaeology, University College Dublin, Ireland
#   
# title-block-published: "Last updated"  
# date: now
# date-format: long
# ---  Output Options ---
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 1000
  cache: true
filters:
 # - ../templates/scholarly-metadata.lua
 # - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/society-for-american-archaeology.csl" # Insert path for the bib-style
reference-section-title: "References Cited" # Required title for the reference list
---

@@

```{r}
#| label: fig-panel-wiki-self-report-numbers
#| fig-cap: "Number of users self-reporting in each category of 'current users' per year"

@@

how_many_people_use_the_wiki <- 
tidy_table_to_plot %>% 
  filter(status != "How many people use the wiki?") %>% 
  mutate(status = str_wrap(status, 26)) %>% 
ggplot() +
  aes(year,
      count,
      fill = status) +
  geom_col() +
  theme_minimal(base_size = small_base_size) +
  xlab("") +
  scale_fill_manual(values = the_thing_palette) +
  theme(legend.position = c(.7,.8),
        legend.key.size = unit(0.25,"line"),
        legend.text = element_text(margin = margin(r = 10, 
                                                   l = 2,
                                                   t = 2,
                                                   b = 2,
                                                   unit = "pt"))) +
  guides(fill=guide_legend(nrow=3,
                           byrow=TRUE),
         )

# set the fig width that we want in the output docx (found this number by trial and error)
out_width <- "6in"
```

```{r}
#| label: fig-panel-wiki-characteristics
#| fig-cap: "A: Total number of edits to each Wiki page (additions and deletions) per year. B: Number of unique editors, as represented by distinct usernames or IP addresses (recorded for editors without usernames), active per year. C: Active periods of individual editors, the black lines connect the date of the first and last edit of all editors that were active for three months or more. D: Typical size of each edit per year, either addition or removal of text, one word is roughly 4-6 bytes. E: Distribution of the numbers of edits per job ad for each year. F: Histogram showing that the majority of editors are only active on the site for less than one year. G: Breakdown of self-reported Wiki user categories for each year."
#| out-width: !expr out_width

@@

```{r}
#| label: fig-show-basic-plots
#| fig-cap: "A: total number of job ads from US institutions posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank. B: Ratio of tenure-track to non-tenure-track positions over time. The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
#| out-width: !expr out_width

@@

```{r}
#| label: fig-show-map-of-hiring-institution-and-cc
#| fig-cap: "A: Frequency of hiring institution by Carnegie classification. B: Inset shows map of the United States showing the count of tenure-track job ads posted by all institutions in each state during 2013 -- 2023"
#| out-width: !expr out_width

@@

```{r}
#| label: fig-show-geo-trends
#| fig-cap: "A: Frequency of locations mentioned in the text of the job ads. B: Popularity of locations in job ads over time for locations that appear in 20 or more ads. Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
#| out-width: !expr out_width

@@

```{r}
#| label: fig-show-metho-trends
#| fig-cap: "A: Frequency of methods mentioned in the text of the job ads. B: Popularity of methods in job ads over time for methods that appear in 10 or more ads. Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
#| out-width: !expr out_width

@@

```{r}
#| label: fig-show-topi-trends
#| fig-cap: "A: Frequency of topics mentioned in the text of the job ads. B: Popularity of topics in job ads over time for topics that appear in 20 or more ads. Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
#| out-width: !expr out_width

@@

```{r}
#| label: fig-show-cooc
#| fig-cap: "Heatmap of topic co-occurrence in job ads."
#| out-width: !expr out_width

@@

```{r}
#| label: fig-requirements-over-time
#| fig-cap: "Changing requirements in job ads over time. 'Count' refers to the number of each item required."
#| out-width: !expr out_width
